---
title: "Intro to R:D2"
output:
  html_document:
    df_print: paged
---
# Set RNotebook for Root Directory
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Desktop/af-werx/data")
```


# Set Working Directory
```{r}
setwd("~/Desktop/af-werx")
```

# Getting working Directory
```{r}
getwd()
```

# creating and setting the working directory
```{r}
main_dir = "~/Desktop/af-werx"
data_dir = paste0(main_dir, "/data")
data_dir
plot_dir = paste0(main_dir, "/plots")
plot_dir
setwd(data_dir)
getwd()
```
# Sample
```{r}
sample(1:100, size = 10)
```

```{r}
sample(state.name, size = 5)
```

# Set Seed to keep same random numbers
```{r}
#set.seed()
sample(state.name, size = 5)
```
# for loop
```{r}
out_list = list()
for(i in 1:15){
  out_list[i] = sqrt(i)
}
out_list
```
# lapply
```{r}
out_list = lapply(1:5, sqrt)
out_list
```

```{r}
state.region
```
# Unique region from state.region
```{r}
unique_region = unique(state.region)
unique_region
```
# convert unique_region as
```{r}
unique_region_list = as.list(unique_region)
unique_region_list
```

```{r}
as.character(unique_region_list[[1]])
```
# since it is a factor we want to transform it to a list 
```{r}
unique_region_list = lapply(unique_region_list, as.character)
unique_region_list
```
# to uppercase
```{r}
toupper(unique_region_list[[1]])
```
# to convert the whole list to upper 
```{r}
unique_region_list = lapply(unique_region_list, toupper)
unique_region_list
```
# create numeric list
```{r}
numeric_list = list(first=-1, second =2, third=3, fourth = -4)

```
# suppose we want only positive numbers
# we can apply the abs() function to the list
# to remove the negatives
```{r}
numeric_list_abs = lapply(numeric_list, abs)
numeric_list_abs
```
# how to replace a pattern w/ Gsub
```{r}
gsub(" ", "_", unique_region_list[[4]])
```
# anoynous function w/ lapply
```{r}
fancy_list = lapply(unique_region_list, function(entry){gsub(" ", "_",tolower(entry))})
fancy_list
```
# Create global functions to use in lapply later
```{r}
toUpperAndUnderscore = function(entry){
  gsub(" ", "_", toupper(entry))
}
#to check that it works
toUpperAndUnderscore("north central")

```
# use above global function in lapply
```{r}
fancy_list = lapply(unique_region_list, toUpperAndUnderscore)
fancy_list
```

# generate a random sequence of numbers
```{r}
sample_vec = sample(1:50, size=50)
sample_vec
```
# Get randomized state names
```{r}
state.name[sample_vec]
```
# check if inputed sample_vec[2] is even
```{r}
if(sample_vec[2] %% 2 == 0){
  TRUE
}else{
  FALSE
}
```
# check if inputed sample_vec[2] is odd
```{r}
IsOdd = function(n){
  if(n %% 2 != 0){
    TRUE
  }else{
    FALSE
  }
}
```
# IsEven Function
```{r}
IsEven = function(n){
  if(n %% 2 == 0){
    TRUE
  }else{
    FALSE
  }
}
```
# use sample_vec in IsEven function and flatten list using unlist
```{r}
logical_list = lapply(sample_vec, IsEven)
logical_vec = unlist(logical_list)
logical_vec
```
# pass logical_vec back into sample_vec to get back index into even_id based on only TRUE values
```{r}
even_id = sample_vec[logical_vec]
even_id
```
# get onlys stat.name with even_ids 
```{r}
state.name[even_id]
```

# try sapply to output vector from list, vector, dataframe
```{r}
logical_vec = sapply(sample_vec, IsEven)
logical_vec 
```

# convert state.x77(state_data) to dataframe(state_df)
```{r}
state_data = state.x77
state_df = as.data.frame(state_data)
class(state_data)
class(state_df)
```
# move rownames to state column
```{r}
rownames(state_df)
state_df$State = rownames(state_df)
##view(state_df)
```
# Get rid of rownames 
```{r}
rownames(state_df) = NULL
##view(state_df)
```
# get vector of classes withing state_df
```{r}
variable_class = sapply(state_df, class)
variable_class
```

# get vector of type within state_df
```{r}
variable_type = sapply(state_df, typeof)
variable_type
```
# install a package
```{r}
#install.packages("PACKAGE_NAME")
#library("PACKAGE_NAME")

```
# get help for package
```{r}
#library(help = "PACKAGE_NAME")
```
# install NYCfligts
```{r}
#install.packages("nycflights13")
#library("nycflights13")
```
# load tidyr table data
```{r}
#load("tidyr_tables.RData")
```
# update packages
```{r}
#tidyverse_update()
```
# load flights table from nycflights13 package
```{r}
flights = nycflights13::flights
```
# load data
```{r}
data(flights)
```
# filter flights from month and year
```{r}
#filter_flights = filter(flights, month == 1, year ==2013)
#filter_flights
```
# nested filter for month, day, and dept_time
```{r}
#filter(filter(flights, month==1 & day == 25), dep_time==15)
```
# example of %in% usages to get several critera of a specific variable
```{r}
#filter(flights, month %in% c(1,3,11) & day == 25)
```
# declare a data frame including NA
```{r}
#NA_df = data.frame( x = c(1, NA, 2), y = c(1,2,3))
```

# filter for NA in x column and x values greater/equal to 1
```{r}
#filter(NA_df, is.na(x) | x>= 1)
```

# filter for nly x values greater/equal to one (w/o NA) 
```{r}
#filter(NA_df, x>= 1)
```
