---
title: "Intro to R_D4"
output: html_notebook
---
# Set RNotebook for Root Directory
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Desktop/af-werx/data")
```


# Set Working Directory
```{r}
setwd("~/Desktop/af-werx")
```

# Getting working Directory
```{r}
getwd()
```

# creating and setting the working directory
```{r}
main_dir = "~/Desktop/af-werx"
data_dir = paste0(main_dir, "/data")
data_dir
plot_dir = paste0(main_dir, "/plots")
plot_dir
setwd(data_dir)
getwd()
```

```{r}
#load("tidyr_tables.RData")
flights = nycflights13::flights

```

```{r}
arrange(flights,year,month,day)
```
# order column data using arrange
```{r}
arrange(flights,year,desc(month),day)
```
# create a dataframe with NA values
```{r}
NA_df = data.frame(x=c(1,NA,2), y = c(1,2,3))
```
# see that NA is at the end
```{r}
arrange(NA_df, desc(x))
```
# select on certain columns
```{r}
select(flights, year, month, day)
```
# select a range of columns
```{r}
select(flights, year:day)
```
# exclude certian columns
```{r}
select(flights, -(year:day))
```
# select only specifi columns that end with a key word
```{r}
select(flights, ends_with("delay"))
```

```{r}
flights_sml = select(flights, year:day, ends_with("delay"), distance, air_time)\
flights_sml
```
# create new/calcutaed fields/columns
```{r}
mutate(flights_sml, gain=arr_delay-dep_delay, speed=distance/air_time * 60)
```
# return only new/calculated columns 
```{r}
transmute(flights_sml, gain=arr_delay-dep_delay, speed=distance/air_time * 60)
```
# useful lag function w/ transmute
```{r}
transmute(flights, arr_delay, lead(arr_delay))
```
# ranking using percent_rank
```{r}
?dplyr::ranking
x <- c(3,1,5,2,2)
row_number(x)
percent_rank(x)
```
# summarise function to get mean w/o NA values
```{r}
summarise(flights, delay=mean(dep_delay, na.rm = T))
```
# group by function
```{r}
by_day = group_by(flights, year, month, day)
by_day
```
# using these together 
```{r}
summarise(by_day, delay = mean(dep_delay, na.rm = T))
```
# group by destination and average by distance and delay
```{r}
by_dest = group_by(flights, dest)

delay = summarise(by_dest, count=n(), dist=mean(distance, na.rm = T), delay=mean(arr_delay, na.rm = T))

delay = filter(delay, count>20, dest != "HNL")

delay
```
# now with pipes!!
```{r}
delay = flights %>% group_by (dest) %>% 
  summarise(count=n(), dist=mean(distance, na.rm = T), delay=mean(arr_delay, na.rm = T))
delay
```
# w/ NA included (another example)
```{r}
flights %>% group_by(year, month, day) %>%
  summarise(mean=mean(dep_delay))
```
# w/o NA included (another example)
```{r}
flights %>% group_by(year, month, day) %>%
  summarise(mean=mean(dep_delay, na.rm = T))
```
# with count function ( = n()) included
```{r}
flights %>% group_by(year, month, day) %>%
  summarise(mean=mean(dep_delay, na.rm = T), n = n())
```
# using count another way (( =n())
```{r}
flights %>% count(day)
```

```{r}
flights %>% group_by(year, month) %>% 
  summarise(first=min(dep_time, na.rm = T), last=max(dep_time, na.rm = T))
```

```{r}
min(flights$dep_time, na.rm = T) #1
max(flights$dep_time, na.rm = T) #2400
```

# find not cancelled flights (where !is.na of dep_time = not cancelled)
```{r}
flights %>% 
  filter(!is.na(dep_time)) %>% 
  group_by(year, month, day) %>% 
  summarise(first = min(dep_time), last = max(dep_time))
```
# using distinct to count unique departure times on year, month, day
```{r}
not_cancelled %>% group_by(year,month,day) %>% summarise(flights_that_take_off=n_distinct(dep_time))
```
# ungroup to regroup as needed
```{r}
not_cancelled %>% ungroup() %>%
  group_by(year,month,day) %>%
  summarise(flights_by_year = n_distinct(dep_time))
  
```
# bad data across two columns 
```{r}
year_country
```

# gather function
```{r}
year_country %>%
  gather("1999","2000", key = "year", value ="cases")
```
# by range from column 2 to column 3
```{r}
year_country %>%
gather(2:3, key = "year", value="cases")
```
# bad data in key column
```{r}
key_value_country
```
# using spread to separate key and value columns
```{r}
key_value_country %>% spread(key=key, value = value)
```
# spread function to spread case and population
```{r}
spread(key_value_country, key=key, value=value)
```
# seperate rate into cases and polation by identifying '/'
```{r}
rate_country %>% 
  separate(rate, into=c("cases", "population"), sep="/")
rate_country
```
# another example of separate usign index of character
```{r}
ex_tables = rate_country %>% 
  separate(year, into=c("century", "year"), sep=2, convert = T)
ex_tables
```
# unit to columns of data (sep = "_" is defualt)
```{r}
ex_tables %>% 
  unite(time, century, year, sep="")
```

