---
title: "Intro to R"
output:
  html_document:
    df_print: paged
---
# Set RNotebook for Root Directory
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Desktop/af-werx/data")
```


# Set Working Directory
```{r}
setwd("~/Desktop/af-werx")
```

# Getting working Directory
```{r}
getwd()
```

# creating and setting the working directory
```{r}
main_dir = "~/Desktop/af-werx"
data_dir = paste0(main_dir, "/data")
data_dir
plot_dir = paste0(main_dir, "/plots")
plot_dir
setwd(data_dir)
getwd()
```

# Read CSV
```{r}
temp_heart_rate = read.csv("temp_heart_rate.csv", header = T, stringsAsFactors = F)
```
# Structure of the Data
```{r}
str(temp_heart_rate)
```

# View the first rows of the data
```{r}
head(temp_heart_rate)
```
# View the last rows of the data
```{r}
tail(temp_heart_rate)
```
# View the whole data
```{r}
#View(temp_heart_rate)
```
# Subset of data
```{r}
temp_heart_subset = temp_heart_rate[1:10,]
write.csv(temp_heart_subset, "temp_heart_subset.csv", row.names = F)
```
# View objects that are in enviroment
```{r}
ls()
```
# Save enviroment
```{r}
save.image(file = "myenv.Rdata")
```
# Remove variable
```{r}
rm("temp_heart_rate")
ls()
```
# Remove all enviroment
```{r}
rm(list=ls())
```
# Loading the Enviroment
```{r}
load("myenv.RData")
```

# CMP
```{r}
CMP = read.csv("ChemicalManufacturingProcess.csv", header = T, stringsAsFactors = F)
str(CMP)
```

# View CMP
```{r}
#View(CMP)
```

# Create CMP Column id Variable
```{r}
column_ids = c(1:4, 14:16)
column_ids
```

# Create CMP Subset with the column_ids vector
```{r}
CMP_Subset = CMP[,column_ids]
str(CMP_Subset)
```

# Get Summary
```{r}
summary(CMP_Subset)
```

# Getting Max
```{r}
maxproccess01 = max(CMP_Subset$ManufacturingProcess01, na.rm = T)
maxproccess01
```

# Checking NAs in Manufacturing Process 01
```{r}
is_na = is.na(CMP_Subset$ManufacturingProcess01)
```
# Verify NA
```{r}
#View(CMP_Subset)
```
# Which
```{r}
which(is_na)
```
# Accessing NA values via subset/index
```{r}
CMP_Subset$ManufacturingProcess01[is_na]
```
# Calculate the Mean
```{r}
mean_process01 = mean(CMP_Subset$ManufacturingProcess01, na.rm = T)
mean_process01
```
# Replace na with mean
```{r}
CMP_Subset$ManufacturingProcess01[is_na] = mean_process01
```

# Calculate MAX again for the whole column
```{r}
maxproccess01 = max(CMP_Subset$ManufacturingProcess01)
maxproccess01
```

# Repeat the process for process 2,3
```{r}
is_na_2 = is.na(CMP_Subset$ManufacturingProcess02)
mean_process02 = mean(CMP_Subset$ManufacturingProcess02, na.rm = T)
CMP_Subset$ManufacturingProcess02[is_na_2] = mean_process02
maxproccess02 = max(CMP_Subset$ManufacturingProcess02)
maxproccess02

is_na_3 = is.na(CMP_Subset$ManufacturingProcess03)
mean_process03 = mean(CMP_Subset$ManufacturingProcess03, na.rm = T)
CMP_Subset$ManufacturingProcess03[is_na_3] = mean_process03
maxproccess03 = max(CMP_Subset$ManufacturingProcess03)
maxproccess03

```
# Mean of Yield
```{r}
yield_mean = mean(CMP$Yield)
yield_mean
```
# Create new Column
```{r}
# if yield is more or equal than the mean then write above_average in new_field
# else write below_average
CMP$new_yield = ifelse(CMP$Yield >= yield_mean, "Above_Average", "Below_Average")
head(CMP[,c("Yield", "new_yield")])
tail(CMP[,c("Yield", "new_yield")])
write.csv(CMP[,c("Yield", "new_yield")], "newyield.csv") # save it to a csv file
```



```{r}
CMP_Subset_variable = colnames(CMP_Subset)

start = 3
end = 6

for(i in start:end){
  print(CMP_Subset_variable[i])
}
```
# Hello World Function
```{r}
PrintHelloWorld = function(name){
  hello = paste0("hello World ", name)
  print(hello)
}

PrintHelloWorld("James")
```
# Get Pi Function
```{r}
GetPi = function(){
  pi = 3.141492
  return(pi)

}
GetPi()
```
# Generic Function to replace dataset's NA with mean
```{r}
ImputeNAWithMean = function(dataset){
  
  for(i in 1:ncol(dataset)){
    is_na = is.na(dataset[,1])
    if(any(is_na)){
      na_ids = which(is_na)
      #mean
      var_mean = mean(dataset[,i], na.rm = T)
      dataset[na_ids,i] = var_mean
      message = paste0("replacing NA with mean:", colnames(dataset)[i])
      print(message)
    }
  }
  return(dataset)
}
#sample dataset inputted into function
CMP_subset = CMP[, c(1:4), 14:16]
CMP_subset_imputed = ImputeNAWithMean(CMP_subset)
#View(CMP_subset_imputed)
```
#sample dataset inputted into function
```{r}
CMP_subset = CMP[, c(1:4), 14:16]
CMP_subset_imputed = ImputeNAWithMean(CMP_subset)
#View(CMP_subset_imputed)

```


