---
title: "Intro to R_D8"
output: html_notebook
---
# import of 
```{r}
CMP = read.csv("ChemicalManufacturingProcess.csv", header = TRUE, stringsAsFactors = FALSE)
```

# ui.r function to build reactive histogram with given variable
```{r}
ui <- fluidPage{
  
  #title of the app
  titlePanel("Histogram"),
  
  #Create a slider input
  sliderInput(inputId = "records", # input id
              label = "Number of records:", # input label
              min = 1,  # min number in slider
              max = 50, # max number in slider 
              value = 30) # default value in slider

  # Select input from a list of inputs.
  selectInput(inputId = "variable", #<­ input ID
              label = "Choose the variable:", #<­ input label 
              choices = as.list(names(CMP)), #<­ input list 
              selected = names(CMP)[1]), #<­ default value of input
              # Render the output as plot.
              plotOutput(outputId = "histPlot")
}
```
# server.r function to build reactive histogram with given variable
```{r}
server <­ function(input, output) { 
  
  # Create a renderPlot function.
  output$histPlot <­ renderPlot({

  # Get the bin width from user and adjust the bin size.
  records <­ seq(min(CMP[[input$variable]]), #<­ min bin value is min value of variable chosen
              max(CMP[[input$variable]]), #<­ max bin value is max value of variable chosen length.out = input$bins + 1) #<­ length is value chosen by the user
 
  # Plot the histogram.
  hist(CMP[[input$variable]], #<­ variable name chosen by user
       breaks = records,  #<­ record width
       col = "salmon", #<­ color of the histogram
       main = input$variable, #<­ title of the histogram 
       xlab = input$variable) #<­ x axis label    
  }) 
}
```
#ui.r for reactive radio button fuction 
```{r}
ui <­ fluidPage(

  
  # Title of the app.
  titlePanel("Table"),

  # Radio button to choose the option. 
  radioButtons("option", #<­ input ID
               "Choose the option", #<­ input 
               label c("Head", "Tail")), #<­ options

  # Numeric inputs.
  numericInput("num", #<­ input ID
               "No of observations", #<­ input label
                value = 10, #<­ default value
                min = 1, #<­ minimum value
                max = nrow(CMP), #<­ maximum value
                step = 1), #<­ increase in step of

  # Ouput function to display.
  tableOutput("cmp_table")
  
)
```
#server.r for reactive radio button fuction 
```{r}
server <­ function(input, output) { 
  
  # Render output table.
  output$cmp_table = renderTable(

  # Check input option to display head or tail.
  if(input$option == 'Head'){
    head(CMP[1:10], input$num) #<­ display head of the data
  } 
  else{
    tail(CMP[1:10], input$num) #<­ display tail of the data
  } 
  )
}
```
# ui.r for scatter and summary outputs
```{r}
ui <­ fluidPage(

  # Title of the app
  titlePanel("Scatter plot and summary of two variables"),
  
  # Select input from a list of inputs
  selectInput(inputId = "variable1",    #<- input1 id
              label = "Choose the first variable:", #<- input1 label
              choices = as.list(names(CMP)), #<- input1 list
              selected = names(CMP)[1]),  #<- default value of input1
  
  # Select input from a list of inputs
  selectInput(inputId = "variable2",  #<- input2 id
              label = "Choose the second variable:", #<- input2 label
              choices = as.list(names(CMP)), #<- input2 list
              selected = names(CMP)[2]),  #<- default value of input2
  
  # Render the output as plot
  plotOutput(outputId = "scatterPlot"),
  
  # Print the summary as text output
  verbatimTextOutput("summary")
)
```
# versio.r for scatter and summary outputs
```{r}
server <­ function(input, output) {

  # Create a reactive expression to select two variables choosen by user
  data <- reactive({
    CMP %>% select(input$variable1, input$variable2)
  })
  
  # Print output scatter plot
  output$scatterPlot <- renderPlot({
    
    plot(data(),          #<- the new data selected
         col = "salmon",  #<- color of the plot
         pch = 16)        #<- plot with filled circles
    
  })
  
  # Print the summary output
  output$summary <- renderPrint({
    
    summary(data())     #<- print the summary of data
    
  })
  }
```
# ui.r for boxplots output
```{r}
  # Title of the app
  titlePanel("Boxplot of the chosen variable"),
  
  # Select input from a list of inputs
  selectInput(inputId = "variable",    #<- input id
              label = "Choose the first variable:", #<- input label
              choices = as.list(names(CMP)), #<- input list
              selected = names(CMP)[1]),  #<- default value of input
  
  # Text input to get the title of the plot
  textInput(inputId = "title",       #<- title input id
            label = "write a title", #<- input label
            value = "boxplot"),      #<- default value

  # Render the output as plot
  plotOutput(outputId = "boxPlot")
```
# version.r for boxplot output
```{r}
# Print output scatter plot
  output$boxPlot <- renderPlot({
    
    boxplot(CMP[input$variable],        #<- variable name
            main = isolate(input$title),#<- title name
            col = "aquamarine3")        #<- color of the plot
    
  })
```
# ui.r for correlation plot
```{r}
# Title of the app
  titlePanel("Correlation plot"),
  
  tags$br(),
  tags$br(),
  
  # Action button to trigger the event
  actionButton(inputId = "submit",  #<- input id
               label = "click me"), #<- input label
  
  tags$br(),
  tags$br(),
  
  # Render the output as plot
  plotOutput(outputId = "corr_plot")
```
#version.r for correlation plot
```{r}
observeEvent(input$submit,   #<- observe the click of action button
               
               # Print output scatter plot
               output$corr_plot <- renderPlot({
                 
                 corrplot(cor(CMP[1:10]))
                 
               })
```
#ui.r for event reactive corrplot
```{r}
# Title of the app
  titlePanel("Correlation plot of choosen variables"),
  
  uiOutput("cor_var"), #<- input list of variables to select multiple variables
  
  # Action button to trigger the event
  actionButton(inputId = "submit",  #<- input id
               label = "Update plot"), #<- input label
  
  # Render the output as plot
  plotOutput(outputId = "corr_plot")
  
```
#ui.r for event reactive corrplot
```{r}
  output$cor_var = renderUI({
    selectInput("cor_var",                         #<- input variable id
                "Choose Variables",                #<- label to the select input
                choices = as.list(names(cmp_sub)), #<- choice list
                multiple = T,                      #<- allow multiple selection
                selected = names(CMP)[1:2])        #<- default values
  })
  
  data <- eventReactive(input$submit, {
    cmp_sub[, input$cor_var]     #<- selected input list
  })
  
  output$corr_plot = renderPlot({
               corrplot(cor(data()))  #,- render plot
             })
```
# ui.r for tabular and layouts 
```{r}

  # Give the title to the app
  titlePanel("Understanding layouts"),
  
  # Start the tab window
  tabsetPanel(
    
    # Tab 1
    tabPanel("Histogram 1",                      #<- tab 1 name
             tags$br(), tags$br(),               #<- html breaks
             tags$h2("Histogram of variable 1"), #<- title of plot
             tags$br(), tags$br(),               #<- html breaks
             
             # Select input from a list of inputs
             selectInput(inputId = "variable1",          #<- input id
                         label = "Choose the variable:", #<- input label
                         choices = as.list(names(CMP)),  #<- input list
                         selected = names(CMP)[1]),      #<- default value of input
             
             plotOutput("histogram1")          #<- Plot output
    ),
    
    # Tab 2
    tabPanel("Histogram 2",                       #<- tab 2 name
             tags$br(), tags$br(),                #<- html breaks
             tags$h2("Histogram of variable 2"),  #<- title of the plot
             tags$br(), tags$br(),                #<- html breaks
             
             # Select input from a list of inputs
             selectInput(inputId = "variable2",          #<- input id
                         label = "Choose the variable:", #<- input label
                         choices = as.list(names(CMP)),  #<- input list
                         selected = names(CMP)[1]),      #<- default value of input
             
             plotOutput("histogram2")            #<- plot output
    ),
    
    # Tab 3
    tabPanel("Relationship between two variables",     #<- tab 3 name
             tags$br(), tags$br(),                     #<- html breaks
             tags$h2("Scatter plot of two variables"), #<- title of the plot
             tags$br(), tags$br(),                     #<- html breaks
             plotOutput("scatter_plot")                #<- plot output
             
             )
    )
  )
```
#version.r for tabular and layouts 
```{r}

  output$histogram1 = renderPlot({
    
    # Plot the histogram
    hist(CMP[[input$variable1]],      #<- variable name chosen by user
         col = "salmon",              #<- color of the histogram
         main = input$variable1,      #<- title of the histogram
         xlab = input$variable1)      #<- x axis label
  })
  
  output$histogram2 = renderPlot({
    
    # Plot the histogram
    hist(CMP[[input$variable2]],      #<- variable name chosen by user
         col = "lightblue",           #<- color of the histogram
         main = input$variable2,      #<- title of the histogram
         xlab = input$variable2)      #<- x axis label
  })

  # Print output scatter plot
  output$scatter_plot <- renderPlot({
    
    plot(CMP[[input$variable1]],      #<- variable 1
         CMP[[input$variable2]],      #<- variable 2
         col = "salmon",              #<- color of the plot
         pch = 16,                    #<- plot with filled circles
         xlab = input$variable1,      #<- x axis label
         ylab = input$variable2       #<- y axis label
         )                    
    
  })
  
```

