---
title: "Intro to R_D5"
output:
  html_document:
    df_print: paged
---
# Set RNotebook for Root Directory
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "~/Desktop/af-werx/data")
```


# Set Working Directory
```{r}
setwd("~/Desktop/af-werx")
```

# Getting working Directory
```{r}
getwd()
```

# creating and setting the working directory
```{r}
main_dir = "~/Desktop/af-werx"
data_dir = paste0(main_dir, "/data")
data_dir
plot_dir = paste0(main_dir, "/plots")
plot_dir
setwd(data_dir)
getwd()
```

```{r}
#load("tidyr_tables.RData")
flights = nycflights13::flights

```
# Static Plot Example
```{r}
plot(c(1,4,6,7,8,11,44,66))
```
# read data from data sample
```{r}
CMP = read.csv("ChemicalManufacturingProcess.csv", header =  T, stringsAsFactors = F)
```
# take a look at the data
```{r}
#View(CMP)
```

```{r}
column_ids = c(1:4,14:16)
column_ids
```
# display column ids and first few variables
```{r}
CMP_subset = CMP[,column_ids]
str(CMP_subset)
```

```{r}
summary(CMP_subset$Yield)
```
# display a simple boxplot
```{r}
boxplot(CMP_subset$Yield)
```
# make it orange and give it a name
```{r}
boxplot(CMP_subset$Yield, col= "orange", main ="Yield Summary")
```
# display color names range 1:6
```{r}
colors()[1:6]
```
 

 # display color help and examples
```{r}
?colors
demo(colors)
```
# get random set of color set
```{r}
set.seed(2)
n_cols = ncol(CMP_subset)
col_sample = sample(colors(), n_cols)
col_sample
```
# use colors in boxplot
```{r}
boxplot(CMP_subset, col = col_sample)
```
# display histogram data without plotting
```{r}
hist(CMP_subset$Yield, plot = F)
```
# plotting histogram with vector of sample colors with x-axis and title changed
```{r}
hist(CMP_subset$Yield, col=col_sample[1:3], xlab = "Yield", main = "Dist. of Yield" )
```
# using par(mfrow = c(x,x)) to designate the the layout of the below histograms
```{r}
par(mfrow = c(1,2))
hist(CMP_subset$BiologicalMaterial01, col=col_sample[2], xlab = "Bio Material 1", main = "Dist. of Bio Material 1")
hist(CMP_subset$BiologicalMaterial02, col=col_sample[3], xlab = "Bio Material 2", main = "Dist. of Bio Material 2")
```
# display three histograms in dataframe view of 1 row, 3  columns
```{r}
par(mfrow = c(1,3))
hist(CMP_subset$BiologicalMaterial01, col=col_sample[2], xlab = "Bio Material 1", main = "Dist. of Bio Material 1")
hist(CMP_subset$BiologicalMaterial02, col=col_sample[3], xlab = "Bio Material 2", main = "Dist. of Bio Material 2")
hist(CMP_subset$BiologicalMaterial03, col=col_sample[4], xlab = "Bio Material 3", main = "Dist. of Bio Material 3")
```
# creat plot with pch title, x/y-axis lables, and pch symbol mod with 2x enlargement, color =steelblue
```{r}
plot(CMP_subset[,2],CMP_subset[,1], xlab = "Bio Material 1", ylab = "Yield", main = "Bio. Material 1 vs Yield",pch = 4, cex = 2, col ="steelblue")
```
# correlation matrix for quick analysis
```{r}
pairs(CMP_subset[,1:4], pch=19, col="steelblue")
```
# install/load  corrlot package
```{r}
#install.packages("corrplot")
library(corrplot)
library(help="corrplot")
```
# chart a corrplot on std deviation of 1
```{r}
CMP_cor = cor(CMP_subset[,1:4])
#View(CMP_cor)
```
# render it with a change in vis method
```{r}
corrplot(CMP_cor, method = "pie")
```
# display mixed corrplot
```{r}
corrplot.mixed(CMP_cor)
```

```{r}
#install.packages("ggplot2")
```

```{r}
library(ggplot2)
```

```{r}
?ggplot2
```

```{r}
ggp1 = ggplot(CMP_subset, aes(x=Yield))
ggp1
```
# Layered geom_histogram with binwidth 0.75, outline/fill color
```{r}
ggp1 = ggp1 + geom_histogram(aes(y= ..density..), binwidth = 0.75, color="steelblue", fill="gray")
```
# layer density on top of histogram
```{r}
ggp1 = ggp1 + geom_density(alpha=.5, color="gray", fill="steelblue")
ggp1  
```
# using labs function to add title and subtitle
```{r}
ggp1 = ggp1 + labs(title="Distrobution", subtitle="Histogram & Density")
ggp1
```

```{r}
ggp2 = ggplot(CMP_subset, aes(x=BiologicalMaterial01, y=Yield))
ggp2
```

```{r}
ggp2 = ggp2 + geom_point()
ggp2
```

```{r}
ggp2 = ggp2 +
  geom_point(color="darkorange")+
  geom_smooth(method = lm)+
  labs(title="Bio. Material 1 vs Yield", subtitle = "Scatterplot with linear fit")
ggp2
```

```{r}
ggtheme1 = theme_bw() + theme(axis.title = element_text(size = 20),
                            axis.text = element_text(size = 16), 
                            plot.title = element_text(size = 25), 
                            plot.subtitle = element_text(size = 18))
```

```{r}
ggp2 = ggp2 + ggtheme1
ggp2
```

